---
title: "NAB spore analyses"
author: "Yiluan Song"
date: "11/18/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE,  warning = FALSE, message = FALSE)

source("code/0_utils.R")
```

# __Explore data__

## Read and clean NAB data
The complete dataset has both pollen and spore count. In this analysis we focus on spores.
```{r read}
# source("code/1_read.R")
nab_df<-read_rds(str_c(.path$nab_clean,"nab_dat.rds")) %>% as_tibble()
nab_df %>% filter(group == "spore")
```


## Clean up taxonomy
We retrieved classification on various levels for aggregation.
```{r taxa}
# source("code/2_taxonomy.R")
taxa_df<-read_rds(str_c(.path$nab_clean,"nab_taxa.rds")) %>% as_tibble()
taxa_df %>%  filter(kingdom == "Fungi")
```

Join.
```{r join}
source ("code/3_join.R")
df
```

## Sampling stations
Meta-data associated with sampling stations.
```{r meta}
source("code/4_meta.R")
meta_df
```

A total of 34 sampling stations distributed across the US. Selected stations (highlighted in red) are visualized for fungal spore phenology.
```{r map, out.width="100%"}
p_map
```

## Time series
```{r calendar, out.width="100%"}
source ("code/5a_calendar.R")
p_calen
```

## Visualize total spore count
```{r total, out.width = "100%"}
source("code/5b_time_series.R")
p_ts
```

## Visualize community composition
We here show the four most dominant families in comparison to the total spore count.
```{r comm, out.width = "100%"}
source("code/5c_composition.R")
p_comm
```
## Get key phenological metrics

## Temporal trends

## Visualize phenology
Focusing on the family _Cladosporiaceae_, we visualize the phenology of spores. This figure shows the inter-annual and intra-annual variations in spore count. Note that we are not following calendar years but rather start counting 60 days after Jan 1. This is because the spore count is generally the lowest around day 60. We hope to make sure each cycle has an increasing and then decreasing trend in spore count.

```{r pheno, out.width = "100%", fig.height=12}
source("code/12_taxa_phenology.R")
p_pheno
```

## Naive way to fit model

If we fit each phenology curve with a double logistic phenology model, the phenology can sometimes be described well. However, there are undesirable behaviours, e.g., overfitting, large variations in model parameters that can hardly be interpreted. This is particularly problematic for sites and years without a single peak in spore count or have large observational noise.

```{r simplefit}
source("code/13_simple_fit.R")
p_simple
```

# __Fit hierarchical Bayesian model__

We adopt a hierarchical Bayesian modeling approach to impose reasonable constraints and prior knowledge on model parameters. We are also able to borrow strength from multiple time series to find the general shape of phenology curve.

## Model structure
Due to the very different shapes of phenology curves among sites, we fit a hierarchical Bayesian model for each site independently. We start with a data model, assuming spore count to have a Poisson distribution with the mean $\lambda$, and $log(\lambda)$ to follow a double logistic curve over a year (Elmore et al., 2012).


$$
\\
y \sim Poisson(\lambda)\\
log(\lambda)= mn + ((mx-mn) + rsu \times  doy) \times(\frac {1}{1+exp((sos-doy)/rsp)}-\frac {1}{1+exp((eos-doy)/rau)})
\\
mn = mn' \times (mn_{max}-mn_{min}) + mn_{min}\\
mx = mx' \times (mx_{max}-mz_{min}) + mx_{min}\\
sos = sos' \times (sos_{max}-sos_{min}) + sos_{min}\\
eos = eos' \times (eos_{max}-eos_{min}) + eos_{min}\\
rsp = rsp' \times (rsp_{max}-rsp_{min}) + rsp_{min}\\
rau = rau' \times (rau_{max}-rau_{min}) + rau_{min}\\
rsu = rsu' \times (rsu_{max}-rsu_{min}) + rsu_{min}\\
$$

Each parameter is scaled to $(0,1)$ and then logit-transformed to the real number line. We assume that the transformed parameters each has a mean and year-specific random effects.
$$
\\
logit (mn') =\mu_{mn}+\epsilon_{mn,t}\\
logit (mx') =\mu_{mx}+\epsilon_{mx,t}\\
logit (sos') =\mu_{sos}+\epsilon_{sos,t}\\
logit (eos') =\mu_{eos}+\epsilon_{eos,t}\\
logit (rsp') =\mu_{rsp}+\epsilon_{rsp,t}\\
logit (rau') =\mu_{rau}+\epsilon_{rau,t}\\
logit (rsu') =\mu_{rsu}+\epsilon_{rsu,t}\\
\\
\epsilon_{mn,t} \sim N(0, \sigma^{2}_{mn})\\
\epsilon_{mx,t} \sim N(0, \sigma^{2}_{mx})\\
\epsilon_{sos,t} \sim N(0, \sigma^{2}_{sos})\\
\epsilon_{eos,t} \sim N(0, \sigma^{2}_{eos})\\
\epsilon_{rsp,t} \sim N(0, \sigma^{2}_{rsp})\\
\epsilon_{rau,t} \sim N(0, \sigma^{2}_{rau})\\
\epsilon_{rsu,t} \sim N(0, \sigma^{2}_{rsu})\\
$$

The priors and hyperparameters are specified below.
$$
\\
\mu_{mn}, \mu_{mx} , \mu_{sos} ,\mu_{eos} , \mu_{rsp} , \mu_{rau}, \mu_{rsu} \sim N(0,1)\\
\\
\sigma^{2}_{mn}, \sigma^{2}_{mx}, \sigma^{2}_{sos}, \sigma^{2}_{eos}, \sigma^{2}_{rsp}, \sigma^{2}_{rau},\sigma^{2}_{rsu}  \sim Gamma(1, 1e^{-5})
$$

$$
\\
mn_{min} = -5, mn_{max}=5\\
mx_{min} = 5, mx_{max}=15\\
sos_{min}=0, sos_{max}=183\\
eos_{min}=183, eos_{max}=365\\
rsp_{min}=1, rsp_{max}=61\\
rau_{min}=1, rau_{max}=61\\
rsu_{min}=-0.01, rsu_{max}=0.01\\
$$

## Fitting model

We build the model and run MCMC with the _nimble_ package.
```{r model_setup}
# source("code/14b_bayes_fit.R")
```

## Check model fit

This figure shows observed phenology in dashed lines and model-fitted phenology in solid lines. Colors show different sites.
```{r fit, out.width="100%", fig.height=30}
source("code/14d_bayes_eval.R")
p_fitline
```

The Pearson correlation between observed and fitted values is `r round(cor(df_pheno_model_all_years_new$count, df_pheno_model_all_years_new$y_pred),4)`.
```{r corr, out.width = "100%", fig.height=15}
p_corr
```

## Inspect model parameters

At first glance, it is not obvious how the model parameters change with sites or year. Colors show different sites.
```{r params_year, out.width = "100%"}
source("code/15_analyze_bayes_param.R")
p_param_year
```

We do see interesting relationships when comparing these phenology model parameters to climatic variables, i.e., mean annual temperature and total annual precipitation. Smoothed line in blue is from generalized linear regression. Colors show different sites. It appears that the phenology parameter-climate relationship is mainly driven by spatial patterns (among sites), rather than temporal patterns (among years).
```{r params_clim, out.width = "100%", fig.height=10}
p_mat
p_tap
```

Note:
* mn: lower bound of phenology curve
* mx: higher bound of phenology curve
* sos: timing of spring inflection point
* eos: timing of autumn inflection point
* rsp: 1 / rate of increase in spring
* rau: 1 / rate of decrease in fall
* rsu: rate of change during summer 

Regression of logit(transformed parameters) against climatic variables support the possible influence of climate on spore phenology.
```{r reg}
glm_outputlist
```

## References
Elmore, A.J., S.M. Guinn, B.J. Minsley and A.D. Richardson (2012): Landscape controls on the timing of spring, autumn, and growing season length in mid-Atlantic forests. - Global Change Biology 18, 656-674.
