---
title: "Processing"
author: "Ruoyu Wu, Yiluan Song"
date: "2025-04-24"
output:
  html_document: default
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE)

# library(phenologyspore)

.fulldata <- T

if (exists(".fulldata") && isTRUE(.fulldata)) {
  .path <- list(
    input = "../alldata/input/",
    intermediate = "../alldata/intermediate/",
    output = "../alldata/output/"
  )
}
```

Read spore data
```{r read}
if (.fulldata) {
  df_spore <- read_rds(str_c(.path$input, "dat_spore.rds"))
}
```

Filter to stations in continental US and specified family
Rename stations and tag number for each station
Pad the date to full year round
```{r process nab data}
if (.fulldata) {
  df_full <- tidy_nab_data(df_spore, f = "Total", n_m = 10, n_yr = 3, rename = T, pad = T)
}
```

Generate and save site meta data for future downloading of daymet data
```{r meta}
if (.fulldata) {
  df_meta <- df_full %>%
    distinct(n, .keep_all = T) %>%
    select(lat, lon, station, city, state, country, id, n)
  write_rds(df_meta, str_c(.path$intermediate, "meta.rds"))
  usethis::use_data(df_meta, overwrite = T)
}
```

Read the offset and insert into the dataset
Linear interpolation and whittaker smoothing
Define spore year
Tag data completeness in each station-sporeyear

```{r process time series}
if (.fulldata) {
  df_ts <- tidy_nab_ts(df_full, path_offset = str_c(.path$intermediate, "fill_smooth_offset.rds"))
}
```

Extract spore season metrics
```{r metrics}
if (.fulldata) {
  df_metrics <- calc_metrics(df_ts)
  write_rds(df_metrics, str_c(.path$intermediate, "metrics.rds"))
  usethis::use_data(df_metrics, overwrite = T)
}
```

Download the climate data
```{r download daymet}
if (.fulldata) {
  df_daymet_raw <- down_daymet_station_date(df_meta, path = str_c(.path$input, "dat_daymet_raw.rds"))
}
```

Calculate MAT, TAP in each spore year
```{r climate summary}
if (.fulldata) {
  df_daymet_annual <- calc_climate_stationspyr(df_daymet_raw = df_daymet_raw, df_spore = df_ts)
  write_rds(df_daymet_annual, str_c(.path$intermediate, "daymet_annual.rds"))
  usethis::use_data(df_daymet_annual, overwrite = T)
}
```

Join MAT, TAP with metrics data
Filter stations for each metric based on the number of years
```{r tidy metric clim}
df_analysis <- tidy_metric_clim(df_metrics, df_daymet_annual, pct = 0.8)
write_rds(df_analysis, str_c(.path$intermediate, "analysis.rds"))
usethis::use_data(df_analysis, overwrite = T)
```
