---
title: "Processing"
author: "Ruoyu Wu, Yiluan Song"
date: "2025-04-24"
output:
  html_document: default
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE)

# library(phenologyspore)

.fulldata <- T

if (exists(".fulldata") && isTRUE(.fulldata)) {
  .path <- list(
    # dat_process = "../alldata/processed/2023-04-25/",
    # ana_fun = "../functions/",
    input = "../alldata/input/",
    intermediate = "../alldata/intermediate/",
    output = "../alldata/output/"
  )
}
```

Read spore data
```{r read}
if (.fulldata) {
  df_spore <- read_rds(str_c(.path$input, "dat_spore.rds"))
}
```

Filter to stations in continental US and specified family
Rename stations and tag number for each station
Pad the date to full year round
```{r filter}
if (.fulldata) {
  df_full <- tidy_nab_data(df_spore, f = "Total", n_m = 10, n_yr = 3, rename = T, pad = T)
}
```

Generate and save site meta data for future downloading of daymet data
```{r meta}
if (.fulldata) {
  df_meta <- df_full %>%
    distinct(n, .keep_all = T) %>%
    select(lat, lon, station, city, state, country, id, n)
  write_rds(df_meta, str_c(.path$intermediate, "meta.rds"))

  # usethis::use_data(df_meta, overwrite = T)
} else {
  df_meta <- read_rds(str_c(.path$intermediate, "meta.rds"))
}
```

Read the offset and insert into the dataset
```{r read offset and merge with smooth data}
if (.fulldata) {
  source(str_c(.path$ana_fun, "read_offset_station.R"))
  df_full <- df_full %>%
    left_join(df_offset, by = c("lat", "lon", "station", "state", "country", "id")) %>%
    rename(lmd_offset = lambda)
}
```

Linear interpolation and whittaker smoothing
```{r smooth}
source(str_c(.path$ana_fun, "tidy_smoothfillwhit_station.R"))
if (.fulldata) {
  df_smooth_100 <- df_full %>%
    tidy_smoothfillwhit_station(n_gap = 14, lambda = 100, column_name = count, new_column_name = count_fillwhit)
}
```

Define spore year
```{r spore year}
source(str_c(.path$ana_fun, "tidy_sporeyr_station.R"))
if (.fulldata) {
  df_sporeyr <- df_smooth_100 %>% tidy_sporeyr_station()
}
```

Tag data completeness in each station-sporeyear
```{r}
source(str_c(.path$ana_fun, "calc_completeness_stationspyr.R"))
if (.fulldata) {
  df_spore_cpltness <- calc_completeness_stationspyr(df_sporeyr, column_name = count_fillwhit)
}
```

Extract spore metrics
```{r metrics}
source(str_c(.path$ana_fun, "calc_metrics_stationspyr.R"))
if (.fulldata) {
  df_metrics <- calc_metrics_stationspyr(df_completeness = df_spore_cpltness, df_raw = df_sporeyr)
  write_rds(df_metrics, str_c(.path$intermediate, "metrics.rds"))
} else {
  df_metrics <- read_rds(str_c(.path$intermediate, "metrics.rds"))
}
```

Gather df_metrics into a long df
```{r}
source(str_c(.path$ana_fun, "tidy_gathermetrics.R"))
df_metrics_long <- tidy_gathermetrics(df_metrics)
```

Assign color palette for 55 stations
```{r color palette}
color_palette <- colorRampPalette(colors = rainbow(55))
df_metrics_long <- df_metrics_long %>% mutate(col = color_palette(55)[n])
```

Download the climate data
```{r download daymet}
source(str_c(.path$ana_fun, "down_daymet_station_date.R"))
if (FALSE) {
  dat_daymet_raw <- down_daymet_station_date(df_meta) %>%
    filter(measurement %in% c("prcp..mm.day.", "tmax..deg.c.", "tmin..deg.c.")) %>%
    select(-tile)
  write_rds(dat_daymet_raw, str_c(.path$input, "dat_daymet_raw.rds"))
} else {
  df_daymet_raw <- read_rds(str_c(.path$input, "dat_daymet_raw.rds"))
}
```

Calculate MAT, TAP in each spore year
```{r climate summary}
source(str_c(.path$ana_fun, "calc_climate_stationspyr.R"))
if (FALSE) {
  df_daymet_annual <- calc_climate_stationspyr(df_daymet_raw, df_sporeyear = df_sporeyr)
  write_rds(df_daymet_annual, str_c(.path$intermediate, "daymet_annual.rds"))
} else {
  df_daymet_annual <- read_rds(str_c(.path$intermediate, "daymet_annual.rds"))
}
```

Join MAT, TAP with metrics data
```{r data}
df_ana_full <- right_join(df_daymet_annual, df_metrics_long, by = c("lat", "lon", "station", "city", "state", "country", "id", "n", "offset", "year_new"))
```

Filter stations for each metric based on the number of years
```{r}
source(str_c(.path$ana_fun, "filt_station_NofYears_metric.R"))
df_ana <- filt_station_NofYears_metric(df_in = df_ana_full, pct = 0.8)
write_rds(df_ana, str_c(.path$intermediate, "analysis.rds"))
```
