---
title: "analysis"
author: "Ruoyu Wu"
date: "2023-09-29"
output:
  html_document: default
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE)
source("~/Github/spore_phenology/analysis/functions/utils.R")
```

# read spore data
```{r read, message=FALSE, warning=FALSE}
df_spore <- read_rds(str_c(.path$dat_process, "spore_dat.rds"))
```

# filter the stations
```{r filter, message=FALSE, warning=FALSE}
source(str_c(.path$ana_fun, "filt_taxa_station_raw_data.R"))
df_siteyear <- filt_taxa_station_raw_data(df_spore, "Total", 10, 3)
```

# rename stations and tag number for each station
```{r message=FALSE, warning=FALSE}
source(str_c(.path$ana_fun, "tidy_rename_number_station.R"))
df_site <- tidy_rename_number_station(df_siteyear)
```

# generate and save site meta data for future downloading of daymet data
```{r meta, eval = F, message=FALSE, warning=FALSE}
df_meta <- df_site %>%
  distinct(n, .keep_all = T) %>%
  dplyr::select(-year, -date, -count)
# write_rds(df_meta, str_c(.path$dat_process, "meta.rds"))
```

# summarize the ecoregion of each station
```{r ecoregion, eval = F, message=FALSE, warning=FALSE}
source(str_c(.path$ana_fun, "summ_ecoregion_station.R"))
df_ecoregion <- summ_ecoregion_station(df_meta)
```

<!-- # plot ecoregion -->
<!-- ```{r plot-ecoregion, eval = F, message=FALSE, warning=FALSE} -->
<!-- source(str_c(.path$ana_fun, "plot_ecoregion_stations.R")) -->
<!-- df_meta <- read_rds(str_c(.path$dat_process, "meta.rds")) -->
<!-- p_ecoregion <- plot_ecoregion_staton(df_in = df_meta) -->
<!-- p_ecoregion -->
<!-- ``` -->

# pre-process spore raw data: pad the date to full year round
```{r full, message=FALSE, warning=FALSE}
source(str_c(.path$ana_fun, "tidy_fulldate_station.R"))
df_full <- tidy_fulldate_station(df_site)
```

# read the offset and insert into the dataset
```{r read offset and merge with smooth data, message=FALSE, warning=FALSE}
source(str_c(.path$ana_fun, "read_offset_station.R"))
df_full <- df_full %>%
  left_join(df_offset, by = c("lat", "lon", "station", "state", "country", "id")) %>% 
  rename(lmd_offset = lambda)
# write_rds(df_full, str_c(.path$dat_process, "dat_spore_fulldate.rds"))
```

# linear interpolation and whittaker smoothing
```{r smooth, message=FALSE, warning=FALSE}
source(str_c(.path$ana_fun, "tidy_smoothfillwhit_station.R"))
df_smooth_100 <- df_full %>%
  tidy_smoothfillwhit_station(n_gap = 14, lambda = 100,column_name = count, new_column_name = count_fillwhit)
```

# define spore year
```{r spore year, message=FALSE, warning=FALSE}
source(str_c(.path$ana_fun, "tidy_sporeyr_station.R"))
df_sporeyr = df_smooth_100 %>% tidy_sporeyr_station()
# write_rds(df_sporeyr, str_c(.path$dat_process, "dat_smooth_100.rds"))
```

# tag data completeness in each station-sporeyear
```{r message=FALSE, warning=FALSE}
source(str_c(.path$ana_fun, "calc_completeness_stationspyr.R"))
df_spore_cpltness <- calc_completeness_stationspyr(df_sporeyr, column_name = count_fillwhit)
```

# extract spore metrics
```{r, warning=FALSE, message=FALSE}
source(str_c(.path$ana_fun, "calc_metrics_stationspyr.R"))
df_metrics <- calc_metrics_stationspyr(df_completeness = df_spore_cpltness, df_raw = df_sporeyr)
```

# gather df_metrics into a long df
```{r}
source(str_c(.path$ana_fun, "tidy_gathermetrics.R"))
df_metrics_long <- tidy_gathermetrics(df_metrics)
```

# assign color palette for 55 stations
```{r color palette}
color_palette <- colorRampPalette(colors = rainbow(55))
df_metrics_long = df_metrics_long %>% mutate(col = color_palette(55)[n])
```

<!-- # download the climate data -->
<!-- ```{r download daymet, eval = F} -->
<!-- df_meta = read_rds(str_c(.path$dat_process, "meta.rds")) -->
<!-- source(str_c(.path$ana_fun, "down_daymet_station_date.R")) -->
<!-- df_daymet_allday <- down_daymet_station_date(df_meta) -->
<!-- write_rds(df_daymet_allday, str_c(.path$dat_process, "dat_daymet_raw.rds")) -->
<!-- ``` -->

# read datmet raw data
```{r, eval = F}
df_daymet_raw <- read_rds(str_c(.path$dat_process, "dat_daymet_raw.rds"))
```

# calculate MAT, TAP in each spore year
```{r, eval = F}
source(str_c(.path$ana_fun, "calc_climate_stationspyr.R"))
df_daymet_annual <- calc_climate_stationspyr(df_daymet_raw, df_sporeyear = df_sporeyr)
write_rds(df_daymet_annual, str_c(.path$dat_process, "dat_daymet_annual.rds"))
```

# read MAT, TAP and join with metrics data
```{r data}
df_daymet_annual <- read_rds(str_c(.path$dat_process, "dat_daymet_annual.rds"))
df_ana_full <- right_join(df_daymet_annual, df_metrics_long, by = c("lat", "lon", "station", "city", "state", "country", "id", "n", "offset", "year_new"))
# write_rds(df_ana_full, str_c(.path$dat_process, "dat_ana_full_as4605.rds"))
# df_ana_short <- right_join(df_daymet_annual, df_metrics, by = c("lat", "lon", "station", "city", "state", "country", "id", "n", "offset", "year_new"))
# write_rds(df_ana_short, str_c(.path$dat_process, "dat_ana_short_as4605.rds"))
```

# filter stations for each metric based on the number of years
```{r}
source(str_c(.path$ana_fun, "filt_station_NofYears_metric.R"))
df_ana <- filt_station_NofYears_metric(df_in = df_ana_full, pct = 0.8)
# write_rds(df_ana, str_c(.path$dat_process, "dat_ana_as4605.rds"))
```